---
title: "Report on Data 1"
output: html_document
---

```{r, echo=FALSE}
#data3 <- read.csv('data/creditcard_original.csv')

data3 <- read.csv('./data/whole data.csv',
                  na.strings = c("", "NA", "?","unknown",'999'))
```

# Analyse exploratoire des donn´ees
## Data Summary

Below is a table showing the number of missing values and the number of unique values for each column in Data 1.
- size: `
```{r, echo=FALSE}
cat(dim(data3))
```

```{r, echo=FALSE}
# Display the table using knitr::kable() and center it
library(knitr)
library(kableExtra)

 # render the summary table
 renderDT({
  temp<- data3 
  number_of_columns <- ncol(temp)

  number_of_rows <- nrow(temp)

  y <- do.call(rbind, lapply(names(temp), FUN=function(x) {
    mean <- ifelse(is.numeric(temp[[x]]), round(mean(temp[[x]], na.rm = TRUE), digits = 2), NA)
    median <- ifelse(is.numeric(temp[[x]]), round(stats::median(temp[[x]], na.rm = TRUE), digits = 2), NA)
    sd <- ifelse(is.numeric(temp[[x]]), round(stats::sd(temp[[x]], na.rm = TRUE), digits = 2), NA)
    variance <- ifelse(is.numeric(temp[[x]]), round(stats::var(temp[[x]], na.rm = TRUE), digits = 2), NA)
    max <- ifelse(is.numeric(temp[[x]]), max(temp[[x]], na.rm = TRUE), NA)
    min <- ifelse(is.numeric(temp[[x]]), min(temp[[x]], na.rm = TRUE), NA)
    IQR <- ifelse(is.numeric(temp[[x]]), IQR(temp[[x]], na.rm =TRUE), NA)
    levels <- length(unique(temp[[x]]))
    c(Column=x,
      Class=class(temp[[x]]),
      Missing = sum(is.na(temp[[x]])),
      Unique_Values = as.integer(levels))

  }))
      x= list(Rows = number_of_rows, Columns = number_of_columns, Types = y)
      DT::datatable(x$Types,
                class = 'cell-border stripe',
                selection=list(mode="multiple", target="row"),
                rownames=FALSE,
                options = list(scrollX = TRUE, autoWidth = FALSE)
                )
    }
  )
```
- initial notes:
  - 3 constant columns (Over18, employeeCount, StandardHours)
  - 1 ID class (EmployeeID)
  - 1 Categorical variable (Education)
  - The rest are numeric variables

```{r, echo=FALSE, message=FALSE}
# Step 1: Remove specified columns
data3 <- data3[, !colnames(data3) %in% c("Over18", "employeeCount", "StandardHours", "EmployeeID")]

# Step 2: Convert specified categorical variable 'Education' to factor
data3$Education <- as.factor(data3$Education)

# Step 3: Identify non-numeric integer columns and convert them to factors
non_numeric_int_cols <- sapply(data3, function(x) !is.integer(x) && !is.numeric(x))
data3[non_numeric_int_cols] <- lapply(data3[non_numeric_int_cols], as.factor)
```
## Class Distribution

This section presents a bar plot showing the count of rows in the data frame `df` for each class (0 and 1).
```{r, echo=FALSE, message=FALSE}
ggplot(data3, aes(x = Attrition)) +
  geom_bar(aes(y = ..count..), fill = "skyblue") +
  geom_text(stat = "count", aes(label = ..count..), 
            vjust = -0.5, size = 5) +  # Adjust vjust for label position
  labs(title = "Frequency of Classes", x = "Class", y = "Frequency") +
  theme_minimal()
```
### relation catagorical variable/churn 
```{r, echo=FALSE, message=FALSE}
# Define the numeric columns that should remain numeric
numeric_columns <- names(data3[sapply(data3, function(x) is.integer(x)||is.numeric(x))])
```
- finding the most important catgeorical varialbes:
```{r, echo=FALSE, message=FALSE,warning=FALSE}
# Set the minimum count threshold for each level
min_count_threshold <- 5

# Function to check if a variable has enough observations per level
has_enough_levels <- function(variable, threshold) {
  tbl <- table(variable)
  return(all(tbl >= threshold))  # Check if all counts meet the threshold
}

# Get the names of categorical variables with enough counts
valid_variables <- data3 %>%
  select(-all_of(c(numeric_columns,'Attrition'))) %>%
  summarise(across(everything(), 
                   ~ has_enough_levels(.x, min_count_threshold))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Has_Enough_Levels") %>%
  filter(Has_Enough_Levels) %>%
  pull(Variable)

# Perform Chi-Square test for each valid categorical variable
importanceCat <- data3 %>%
  select(all_of(valid_variables)) %>%  # Select only valid variables
  summarise(across(everything(), 
    ~ chisq.test(table(.x, data3$Attrition))$p.value)) %>%  # Replace 'y' with your binary target
  pivot_longer(cols = everything(), 
    names_to = "Variable", values_to = "P_Value") %>%
  arrange(P_Value)  # Sort by p-value

# Render the DataTable
renderDataTable(
  importanceCat,
  options = list(
    pageLength = 15,  # Show 15 rows per page
    autoWidth = TRUE  # Automatically adjust column widths
  ),
  rownames = FALSE  # Do not show row numbers
)
```

#### Plotting the top 3 
```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)

# Assuming you've already calculated 'importanceCat' with p-values from Chi-Square tests

# Get the top 3 variables based on their p-values
top3_variables <- importanceCat %>%
  top_n(-3, P_Value) %>%  # Select the top 3 variables with the smallest p-values
  pull(Variable)

# Loop through each of the top 3 variables and create a bar plot
for (var in top3_variables) {
  plot_data <- data3 %>%
    group_by(!!sym(var), Attrition) %>%
    summarise(count = n()) %>%
    ungroup()
  
  # Create a bar plot for the variable
  p <- ggplot(plot_data, aes(x = !!sym(var), y = count, fill = factor(Attrition))) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Distribution of", var, "by Churn Status"),
         x = var,
         y = "Count",
         fill = "Churn Status") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)  # Ensure each plot is printed during the loop
}
```
#### Plotting the top 3 rates 
```{r, echo=FALSE, message=FALSE}

library(ggplot2)
library(dplyr)
library(tidyr)

# Assuming you've already calculated 'importanceCat' with p-values from Chi-Square tests

# Get the top 3 variables based on their p-values
top3_variables <- importanceCat %>%
  top_n(-3, P_Value) %>%  # Select the top 3 variables with the smallest p-values
  pull(Variable)

# Loop through each of the top 3 variables and create a churn rate plot
for (var in top3_variables) {
  plot_data <- data3 %>%
    group_by(!!sym(var), Attrition) %>%
    summarise(count = n()) %>%
    group_by(!!sym(var)) %>%
    mutate(churn_rate = count / sum(count)) %>%  # Calculate churn rate
    filter(Attrition=="Yes") %>%  # Keep only the churn rate (where y == 1)
    ungroup()
  
  # Create a bar plot showing churn rate
  p <- ggplot(plot_data, aes(x = !!sym(var), y = churn_rate, fill = factor(Attrition))) +
    geom_bar(stat = "identity", position = "dodge", fill = "steelblue") +
    labs(title = paste("Churn Rate by", var),
         x = var,
         y = "Churn Rate",
         fill = "Churn Status") +
    scale_y_continuous(labels = scales::percent_format()) +  # Show y-axis as percentage
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(p)  # Ensure each plot is printed during the loop
}
```

### relation numerical variable/churn 
- this are the correlations between each numerical variable and the class
```{r, echo=FALSE, message=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)                  # For pivot_longer
library(ggplot2)
library(DT)
# Calculate correlations between numerical variables and the class
importance <- data3 %>%
  # Ensure 'y' is numeric (assuming 'y' is a column in data3)
  mutate(Attrition = as.numeric(as.factor(Attrition))) %>%
  # Apply cor() only to numeric columns, calculating correlation with 'y'
  summarise(across(where(is.numeric), ~ abs(cor(.x, Attrition, use = "complete.obs")))) %>%
  # Pivot to long format for better readability and sorting
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Correlation") %>%
  # Remove the correlation of 'y' with itself
  filter(Variable != "Attrition") %>%
  # Sort the variables by correlation in descending order
  arrange(desc(Correlation))

# Display the correlation data using DT
renderDataTable(
  importance,
  options = list(
    pageLength = 15,  # Show 5 rows per page
    autoWidth = TRUE  # Automatically adjust column widths
  ),
  rownames = FALSE  # Do not show row numbers
)
```
#### plotting the biggest 3 count
```{r, echo=FALSE, message=FALSE,warning=FALSE}
top_variables <- importance %>%
   top_n(Correlation, n = 3) %>%
  pull(Variable)

# Reshape data for ggplot
df_long <- data3 %>%
  select(all_of(top_variables), Attrition) %>%
  pivot_longer(cols = all_of(top_variables), names_to = "Variable", values_to = "Value")

# Create separate plots for each variable
for (variable in top_variables) {
  plot<-ggplot(df_long %>% filter(Variable == variable), aes(x = Value, fill = as.factor(Attrition))) +
    geom_histogram(alpha = 0.5, drop = TRUE) +
    labs(title = paste("Histogram of", variable, "by Class"),
         x = "Value",
         y = "Count",
         fill = "Class") +
    #scale_y_log10() +  # Apply logarithmic scale to the y-axis
    theme_minimal() 

    print(plot)
    }

```
#### Correlation matrix
```{r, echo=FALSE, message=FALSE,warning=FALSE}

    # Select only numeric variables from the dataset
    numeric_vars <- data3 %>% select_if(is.numeric)

    # Ensure there are at least two numeric columns to compute correlation
        # Calculate the correlation matrix
        correlation_matrix <- cor(numeric_vars, use = "complete.obs")

        # Melt the correlation matrix into long format for ggplot
        corr_melted <- reshape2::melt(correlation_matrix)

        # Create the correlation matrix heatmap
        corr_plot <- ggplot(corr_melted, aes(Var1, Var2, fill = value)) +
            geom_tile(color = "white") +
            scale_fill_gradient2(low = "red", high = "blue", mid = "white",
                                 midpoint = 0, limit = c(-1, 1), space = "Lab",
                                 name = "Correlation") +
            #theme_minimal() +
            theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                             size = 12, hjust = 1),
                  axis.text.y = element_text(size = 12)) +
            coord_fixed() +
            ggtitle("Correlation Matrix") +
            theme(plot.title = element_text(face = "bold", color = "#2E8B57", size = 14, hjust = 0.5))
    corr_plot
```
## Conclusion:
- blah blah blah
# Prediction de churn
- we are just going to use v2, v4, v8 and v11 since the dataset is very large
