---
title: "Report on Data 1"
output: html_document
---

```{r, echo=FALSE}
#data1 <- read.csv('data/creditcard_original.csv')
data1 <- read.csv('data/creditcard.csv')
```
# Analyse exploratoire des donn´ees
## Data Summary

Below is a table showing the number of missing values and the number of unique values for each column in Data 1.
- size: `
```{r, echo=FALSE}
cat(dim(data1))
```

```{r, echo=FALSE}
# Display the table using knitr::kable() and center it
library(knitr)
library(kableExtra)

 # render the summary table
 temp<- data1 
 # Table logic
 renderDT({
  number_of_columns <- ncol(temp)

  number_of_rows <- nrow(temp)

  y <- do.call(rbind, lapply(names(temp), FUN=function(x) {
    mean <- ifelse(is.numeric(temp[[x]]), round(mean(temp[[x]], na.rm = TRUE), digits = 2), NA)
    median <- ifelse(is.numeric(temp[[x]]), round(stats::median(temp[[x]], na.rm = TRUE), digits = 2), NA)
    sd <- ifelse(is.numeric(temp[[x]]), round(stats::sd(temp[[x]], na.rm = TRUE), digits = 2), NA)
    variance <- ifelse(is.numeric(temp[[x]]), round(stats::var(temp[[x]], na.rm = TRUE), digits = 2), NA)
    max <- ifelse(is.numeric(temp[[x]]), max(temp[[x]], na.rm = TRUE), NA)
    min <- ifelse(is.numeric(temp[[x]]), min(temp[[x]], na.rm = TRUE), NA)
    IQR <- ifelse(is.numeric(temp[[x]]), IQR(temp[[x]], na.rm =TRUE), NA)
    levels <- length(unique(temp[[x]]))
    c(Column=x,
      Class=class(temp[[x]]),
      Missing = sum(is.na(temp[[x]])),
      Min = min,
      Median = median,
      Max = max,
      Mean = mean,
      SD = sd,
      Variance = variance,
      Unique_Values = as.integer(levels))

  }))
      x= list(Rows = number_of_rows, Columns = number_of_columns, Types = y)
      DT::datatable(x$Types,
                class = 'cell-border stripe',
                selection=list(mode="multiple", target="row"),
                rownames=FALSE,
                options = list(scrollX = TRUE, autoWidth = FALSE)
                )
    })
```
## Class Distribution

This section presents a bar plot showing the count of rows in the data frame `df` for each class (0 and 1).

```{r, echo=FALSE, message=FALSE}
# Load ggplot2
library(ggplot2)

# Create the plot with text labels
ggplot(data1, aes(x = Class)) +
  geom_bar(aes(y = ..count..), fill = "skyblue") +
  geom_text(stat = "count", aes(label = ..count..), 
            vjust = -0.5, size = 5) +  # Adjust vjust for label position
  labs(title = "Frequency of Classes", x = "Class", y = "Frequency") +
  theme_minimal()
```
### relation variable categorial/churn
- there is no catgeorical vriables
### relation categorical variable churn 
- this are the correlations between each numerical variable and the class
```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(tidyr)                  # For pivot_longer
library(ggplot2)
library(DT)
# Calculate correlations between numerical variables and the class
importance <- data1 %>%
  mutate(Class = as.numeric(as.factor(Class))) %>%
  summarise(across(starts_with("V"), ~ abs(cor(.x, Class)))) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Correlation") %>%
  arrange(desc(Correlation))

# Display the correlation data using DT
renderDataTable(
  importance,
  options = list(
    pageLength = 15,  # Show 5 rows per page
    autoWidth = TRUE  # Automatically adjust column widths
  ),
  rownames = FALSE  # Do not show row numbers
)
```
#### plotting the biggest 3 
```{r, echo=FALSE, message=FALSE}
top_variables <- importance %>%
  top_n(3, Correlation) %>%
  pull(Variable)

top_variables
# Reshape data for ggplot
df_long <- data1 %>%
  select(all_of(top_variables), Class) %>%
  pivot_longer(cols = all_of(top_variables), names_to = "Variable", values_to = "Value")

# Create histograms for top correlated variables colored by class
# Filter out non-positive values
ggplot(df_long, aes(x = Value, fill = as.factor(Class))) +
  geom_histogram(
                 alpha = 0.5, 
                 drop=T) +
  facet_wrap(~ Variable) +
  labs(title = "Histograms of Top 3 Correlated Variables by Class",
       x = "Value",
       y = "Count",
       fill = "Class") +
  scale_y_log10() +  # Apply logarithmic scale to the y-axis
  theme_minimal()

```
#### plotting the least 3 
```{r, echo=FALSE, message=FALSE}

top_variables <- importance %>%
   slice_min(order_by = Correlation, n = 3) %>%
  pull(Variable)

top_variables
# Reshape data for ggplot
df_long <- data1 %>%
  select(all_of(top_variables), Class) %>%
  pivot_longer(cols = all_of(top_variables), names_to = "Variable", values_to = "Value")

# Create histograms for top correlated variables colored by class
# Filter out non-positive values
ggplot(df_long, aes(x = Value, fill = as.factor(Class))) +
  geom_histogram(
                 alpha = 0.5, 
                 drop=T) +
  facet_wrap(~ Variable) +
  labs(title = "Histograms of Least 3 Correlated Variables by Class",
       x = "Value",
       y = "Count",
       fill = "Class") +
  scale_y_log10() +  # Apply logarithmic scale to the y-axis
  theme_minimal()

```
### Correlation matrix
```{r, echo=FALSE, message=FALSE}

    # Select only numeric variables from the dataset
    numeric_vars <- data1 %>% select_if(is.numeric)

    # Ensure there are at least two numeric columns to compute correlation
        # Calculate the correlation matrix
        correlation_matrix <- cor(numeric_vars, use = "complete.obs")

        # Melt the correlation matrix into long format for ggplot
        corr_melted <- reshape2::melt(correlation_matrix)

        # Create the correlation matrix heatmap
        corr_plot <- ggplot(corr_melted, aes(Var1, Var2, fill = value)) +
            geom_tile(color = "white") +
            scale_fill_gradient2(low = "red", high = "blue", mid = "white",
                                 midpoint = 0, limit = c(-1, 1), space = "Lab",
                                 name = "Correlation") +
            #theme_minimal() +
            theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                             size = 12, hjust = 1),
                  axis.text.y = element_text(size = 12)) +
            coord_fixed() +
            ggtitle("Correlation Matrix") +
            theme(plot.title = element_text(face = "bold", color = "#2E8B57", size = 14, hjust = 0.5))
    corr_plot
```
## Conclusion:
- blah blah blah
# Prediction de churn
- we are just going to use v2, v4, v8 and v11 since the dataset is very large
